---
title: 'Enterprise Continuous Regression & Stacking Ensemble Report'
author: 'Russ Conte'
date: '`r Sys.Date()`'
format:
  html:
    theme: cosmo
    toc: true
    toc-depth: 2
---

```{r setup, include=FALSE}
pipeline <- readRDS('/Users/russconte/Library/Mobile Documents/com~apple~CloudDocs/Documents/Machine Learning templates in R/NumericEnsembles/NumericEnsembles/numeric_pipeline_data_cache.rds')
```

## 1. Executive Summary
The competitive parallelized regression ensemble pipeline successfully cross-validated models across the continuous space.
The chosen champion stacking model layout is **`r pipeline$performance_report$Model[1]`** achieving an out-of-sample RMSE of **`r pipeline$performance_report$Testing_RMSE[1]`**.

## 2. Predictor Variable Properties & Insights
```{r show-dict, echo=FALSE}
knitr::kable(pipeline$exploratory_insights, caption = 'Automated Pipeline Exploratory Insights Matrix')
```

## 3. Algorithm Standings Leaderboard
```{r show-leaderboard, echo=FALSE}
knitr::kable(head(pipeline$performance_report, 15), caption = 'Team of Rivals Regression Standings Leaderboard')
```
