---
output: github_document
---

```{r, Initial setting, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# NumericEnsembles

The goal of NumericEnsembles is to automatically build highly optimized ensembles of complete solutions where the target column is continuous numeric.

## The story of NumericEnsembles

My name is Russ Conte, and I have worked for many years with multi-million dollar accounts for multi-billion dollar customers, mainly as a recruiter. One of the most common results I have seen are companies who do not use their data to get the best return for their investment to get the data. I had the insight about how to build ensembles of solutions on Saturday, October 22, 2022 at 4:58 pm. The original ensembles solution has been improved many times, and NumericEnsembles is one of 16 ensembles solutions I am building. The total list is:

1.  `NumericEnsembles`

2.  `ClassificationEnsembles`

3.  `LogisticEnsembles`

4.  `ForecastingEnsembles`

5.  `ClusteringEnsembles`

6.  `SurvivalEnsembles`

7.  `TextEnsembles`

8.  `CountingEnsembles`

9.  `SeverityEnsembles`

10. `MultiLabelEnsembles`

11. `NetworkEnsembles`

12. `SpatialEnsembles`

13. `SurveyEnsembles`

14. `LongitudinalEnsembles`

15. `CrossSectionalEnsembles`

16. `CohortEnsembles`

## Installation

You can install the development version of NumericEnsembles from [GitHub](https://github.com/) with:

```{r Install NumericEnsembles}
# library(pak)
# pak::pkg_install("InfiniteCuriosity/NumericEnsembles")
```

## Pipelines: The best way to get results from all the ensembles packages

All 13 ensembles packages work best if you start by building a pipeline first. A pipeline combines all the results (tables, plots, models, and metadata) into one structured asset which you can print, plot, predict, export, save, and much more.

## Track 1: The Express Track (Quick Start)

The Express Track allows you to test your installation instantly using rapid cross-validation configurations and automated synthetic data generations:

```{r NumericEnsembles quick start}
library(NumericEnsembles)

# Using internal demo data generator as an express validation run
Concrete_express_pipeline <- NumericEnsemblesDemo()
```

## Track 2: The Fast (but not as fast as Express) track

This example features facet colors, column colors and stratify colors. You can use these features in many other data sets.

```{r Fast_but_not_express_track}
library(NumericEnsembles)
Insurance <- NumericEnsembles::Insurance
Insurance_pipeline <- Numeric(dataset = Insurance, target_col = 'charges', facet_col = 'sex', color_col = 'smoker', stratify_col = 'region', palette_style = "modern", config = NumericEnsemblesFastConfig(), verbose = TRUE)
print(Insurance_pipeline)
Insurance_pipeline$plots # plots all in one command
```

## Track 3: The Institutional Track (Professional Production, this example will have a lower root mean squared error than an article in Nature for the exact same data set)

For this next example we will be using the `Concrete` data set, and achieving a lower root mean squared error (RMSE) than this article in Nature on the same data set: <https://www.nature.com/articles/s41598-024-69616-9>. The article shows a lowest RMSE of 5.11, NumericEnsembles will get a best RMSE that is lower than 5.11, and you will be able to verify the result.

For enterprise-grade model deployments, you can decouple hyperparameter states from your execution tracks using the complete `NumericEnsemblesConfig()` matrix. This path showcases advanced feature transformations (including `YeoJohnson`power-scaling), high-leverage data outlier filtering via Cook's Distance, and rigorous multi-model hyperparameter tuning grids:

```{r Professional Production}
library(NumericEnsembles)

# 1. Establish custom, comprehensive hyperparameter tuning grids
custom_glmnet_grid <- expand.grid(
  alpha  = seq(0, 1, length = 5),
  lambda = seq(0.001, 0.2, length = 10)
)

custom_rf_grid <- expand.grid(
  mtry = c(2, 4, 6, 8)
)

# 2. Build the fine-grained execution configuration matrix
institutional_config <- NumericEnsemblesConfig(
  cv_folds        = 10,       # Rigid 10-fold cross-validation
  train_pct       = 0.80,     # 80/20 train-test population split
  vif_threshold   = 10.0,     # Strict multicollinearity screening
  cooks_threshold = 2.0,      # Prune high-leverage outliers over 2 * (4/n)
  transform_steps = c("nzv", "medianImpute", "center", "scale", "YeoJohnson"),
  glmnet_grid     = custom_glmnet_grid,
  rf_grid         = custom_rf_grid,
  svm_tune_length = 10,
  pcr_tune_length = 10
)

# 3. Execute the concurrent machine learning rival engine
Concrete_pipeline <- Numeric(
  dataset       = Concrete[1:1000, ], 
  target_col    = 'Strength', 
  palette_style = "modern", 
  config        = institutional_config, 
  verbose       = TRUE
)

# 4 Verify five best results
Concrete_pipeline$performance_report[1:5, ]
```

## Print the Summary Profiles from any NumericEnsembles Pipeline

When you call `print()` on your pipeline object, it outputs a multi-profile metadata evaluation series:

1.  **Baseline Sample Head & Population Description:** Immediate tracking of your baseline raw data structures.

2.  **Structural Data Dictionary:** Maps column classes, missing value counts, and unique value frequencies.

3.  **Automated Exploratory Summary Insights:** Granular tracking of feature anomalies, calculating exact skewness coefficients, IQR outlier counts, and emitting specific operational text insights.

4.  **Multicollinearity Threshold Audit Report:** A complete breakdown of columns evaluated, empirical Variance Inflation Factors (VIF), and selection status ("Kept" vs "Dropped").

5.  **Ensemble Performance Leaderboard Evaluation:** Multi-engine rankings sorted by Testing RMSE, complete with MAE, Adjusted R2, prediction variance, and run durations.

6.  **Automated Residual Diagnostic Leaderboard:** Runs validation scans checking residuals for normality (Shapiro-Wilk), homoscedasticity (Spearman), and error independence (Durbin-Watson).

## Generate Diagnostic Plots from any NumericEnsembles Pipeline

You can interact with your visual diagnostics package in two standard ways:

```{r Generate_diagnostic_plots}
# plot(Concrete_pipeline)  # Sequentially render plots to your active device window
Concrete_pipeline$plots  # Direct programmatic access to specific ggplot2 objects
```

The professional visual diagnostics portfolio includes:

- **Histograms:** Continuous feature density and population distribution panels.

- **Box Plots:** Predictor range distribution quantiles and scale profiles.

- **Correlation Heatmap:** Multi-feature linear explanatory predictor correlation matrices.

- **Scatter Analysis Matrix:** Individual regression line mappings matching the target column against features.

- **Performance Metrics & KPIs:** Horizontal ranking of cross-validated architectures showcasing explicit **95% predictive confidence intervals** across RMSE, MAE, and Adjusted R2.

- **Generalization Risks & Structural Bias:** Mappings tracking overfitting ratios and model directionality bias.

- **Bias-Variance Space:** Joint coordinate mapping of directional model bias vs empirical variance relative to an absolute ideal vector origin.

- **Kolmogorov-Smirnov Test Mappings:** Charting distribution alignment p-values across your candidate algorithms.

- **Cook's Distance Leverage Timeline:** A standalone segment tracker charting raw outlier boundaries.

## Predicting on New Data

Programmatic predictions can be generated instantly using the pipeline's optimized S3 method wrapper, utilizing the absolute top-performing champion model architecture:

```{r Predicting on new data}
prospective_data <- Concrete[1001:1030, ]
Pipeline_predictions <- predict(object = Concrete_pipeline, newdata = prospective_data, model_name = "best")
```

For industrial workloads, use `predict_production()` to automatically obtain point projections alongside matching row-level 95% upper and lower assurance boundaries for the top 3 champion models:

```{r Predict production}
Production_report <- predict_production(object = Concrete_pipeline, newdata = prospective_data)
Production_report
```

## Render the Automated Executive Report

You can compile and render a standalone, professional corporate report from your pipeline instantly. This function extracts your metadata matrix and generates a polished, executive summary document utilizing high-speed local Quarto compilation:

```{r Automated Executive Report}
# Compiles a polished HTML document matching your chosen palette style
RenderExecutiveReport(pipeline_object = Concrete_pipeline, output_directory = getwd())
```

## Professional features of NumericEnsembles

- Fine-grained hyperparameters handling via `NumericEnsemblesConfig()` and rapid cross-validation configurations with `NumericEnsemblesFastConfig()`.

- Error-bound verifications delivering explicit 95% predictive confidence intervals across all primary KPIs.

- 7 professional curated test sample datasets available at the [EnsemblesData Repository](https://github.com/InfiniteCuriosity/EnsemblesData/tree/main/NumericEnsembles).

- Dedicated I/O pipelines for seamless asset transportation via `save_pipeline()`, `load_pipeline()`, and `ExportNumericResults()`.

- Never calls any large language models (LLMs). This is a completely LLM-free, high-performance local algorithmic solution.
