---
title: "Getting Started with NumericEnsembles: Advanced Continuous Stacking Optimization"
author: "Russ Conte"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with NumericEnsembles}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r Getting started, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

**Introduction**:

Ordinary regression analysis often forces analysts to commit to a single structural paradigm—such as Ordinary Least Squares (OLS) or a standard tree. However, real-world data contains complex surface topologies, multi-collinearity constraints, and variable interactions that single models frequently misinterpret.

The `NumericEnsembles` package answers this problem by establishing a highly parallelized, production-grade **Team of Rivals** framework. It trains 17 independent base machine learning architectures simultaneously, feeds their out-of-sample prediction arrays into 5 specialized stacking meta-learners, and evaluates all possible pairwise hybrid blends to construct an optimized ensemble consensus model.

## 1. The Core Stacking & Blending Architecture

To maximize predictive generalization, `NumericEnsembles` leverages three progressive layers of statistical combinations:

### Layer 1: The 17 Base Machine Learning Rivals

The core engine concurrency loop estimates a diverse collection of algorithms spanning linear, tree-based, regularized, and neural network structures:

- **Linear & Component Spaces:** OLS Linear Regression, Generalized Linear Models (GLM), Principal Component Regression (PCR), and Multivariate Adaptive Regression Splines (MARS).

- **Regularization Matrix Containers:** Ridge, Lasso, and ElasticNet regressions optimized via elasticnet/glmnet grids.

- **Tree & Advanced Ensembles:** Bagged Trees, Bagged MARS, Recursive Partitioning (RPart), Cubist Rule-Based Models, Random Forests, Quantile Random Forests, and Conditional Inference Forests.

- **Neural Networks & Splines:** Bayesian Regularized Neural Networks (BRNN), standard Single-Layer Feed-Forward Neural Networks, and Model-Averaged Neural Networks.

### Layer 2: Stacking Meta-Learners

The out-of-sample training predictions of all 17 models are isolated into a primary meta-feature dataframe. This grid is used to train five structural stacking meta-learners—a Baseline Meta-GLM, an Optimized Meta-Elastic Net, a Smooth Non-linear Meta-Boosted GAM, a Meta-Random Forest, and a Meta-SVM—to learn exactly how to weight models based on contextual predictor densities.

### Layer 3: Pairwise Blended Consensus Combinations

The package evaluates all possible pairwise variations (C(17,2)=136 model combinations) using an out-of-sample validation split. It ranks every base model, meta-learner, and blended combination by Root Mean Squared Error (RMSE) to discover the champion configuration.

## 2. End-to-End Operational Walkthrough

Below is an end-to-end workflow demonstrating how to execute the multi-core pipeline over continuous variables and interpret its objects.

### Initialize the Performance Dataset

We will demonstrate using a simulated economic registry featuring continuous indicators and an underlying multi-collinear structure:

```{r Initialize the Performance Dataset}
library(NumericEnsembles)

set.seed(42)
n_samples <- 400

# Simulate continuous features
housing_index <- runif(n_samples, min = 80, max = 150)
unemploy_rate <- rnorm(n_samples, mean = 5.5, sd = 1.2)
# Introduce explicit collinearity to trigger the VIF filter
inflation_prox <- (0.45 * unemploy_rate) + rnorm(n_samples, mean = 2, sd = 0.2)

# Construct a continuous target variable
gdp_growth <- 2.5 + (0.04 * housing_index) - (0.35 * unemploy_rate) + rnorm(n_samples, sd = 0.5)

economic_registry <- data.frame(
  GDP_Growth      = gdp_growth,
  Housing_Index   = housing_index,
  Unemployment    = unemploy_rate,
  Inflation_Proxy = inflation_prox
)
```

### Deploying the Multi-Core Pipeline

We generate an isolated configuration parameter matrix using `NumericEnsemblesConfig()`. This decouples operational arguments (such as cross-validation structures, training percentages, and maximum allowable Variance Inflation Factor limits) from the primary wrap runner before executing the core `Numeric()` engine:

```{r Multi-core pipeline}
# Initialize the decoupled operational parameters matrix 

custom_config <- NumericEnsemblesConfig(
  cv_folds      = 5,          # 5-Fold Cross-Validation per candidate
  train_pct     = 0.70,       # 70% Training / 30% Testing Evaluation Split
  vif_threshold = 5           # Max allowed VIF multiplier boundary
)

# Execute the complete parallel optimization backend using our configuration asset
trained_numeric_ensemble <- Numeric(
  dataset       = economic_registry,
  target_col    = "GDP_Growth",
  palette_style = "modern",   # High-contrast deep navy/teal theme layout
  config        = custom_config,
  verbose       = TRUE
)
```

## 3. Interacting with S3 Objects and Methods

`NumericEnsembles` contains custom S3 methods designed to inspect performance reporting profiles and handle visual analytics arrays cleanly.

### Executive Summary Report

Passing your output object straight into the standard `print()` method prints your continuous evaluation portfolio, showing your training durations, overfitting ratios, and out-of-sample validation leaderboards:

```{r print the methods}
print(trained_numeric_ensemble)
```

### Multidimensional Performance Dashboards

Calling the custom `plot()` method loops through an exhaustive visual audit portfolio, rendering your feature histograms, range boxplots, correlation space, and your core performance KPI panels (displaying RMSE, MAE, and Adjusted R2 with complete asymptotic 95% Confidence Intervals side-by-side):

```{r Visual evaluation}
# Paces through your visual evaluation windows interactively
plot(trained_numeric_ensemble, pace_output = FALSE)
```

## 4. Deploying Scoring Inferences over Untrained Records

To map un-indexed prospects or score incoming continuous spreadsheets, pass your trained pipeline object, the `newdata`data frame, and your chosen leaderboard model identifier directly into the S3 `predict()` wrapper method.

Setting `model_name = "best"` automatically selects the absolute top-performing champion architecture from your validation trials:

```{r Establish a raw out-of-sample row profile using the RandomForest model}
library(NumericEnsembles)
Concrete <- NumericEnsembles::Concrete
Concrete_pipeline1 <- Numeric(dataset = Concrete[1:1000, ], target_col = 'Strength', facet_col = "", color_col = "", stratify_col = "", palette_style = "modern", config = NumericEnsemblesFastConfig(), verbose = TRUE)
Concrete_predictions <- predict(object = Concrete_pipeline1$base_models$RandomForest, newdata = Concrete[1001:1010, ])
Concrete_predictions
```

## 5. Disk Serialization Protocols

For production deployments, `NumericEnsembles` features fast disk serialization utilities to save and re-hydrate your entire modeling workspace environment cleanly without rebuilding your data models from scratch. Un-comment the second line of each set if you want to us these functions.

```{r Disk Serialization Protocols}
# Serialize and compress your entire environment cache to a local file
# save_pipeline(Concrete_pipeline1, "Concrete_pipeline1.rds")

# Re-hydrate your workspace cleanly in a fresh session instantly
# rehydrated_pipeline <- load_pipeline("Concrete_pipeline1.rds")
```

The compilation loop will cleanly parse every single markdown chunk and embed a pristine, production-grade continuous regression manual directly inside your package's documentation library!
