---
title: "Predicting on Unseen Production Data with Strength of Concrete"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Predicting on Unseen Production Data with Strength of Concrete}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  warning = FALSE,
  message = FALSE
)
```

## Introduction

The ultimate validation of a predictive machine learning asset is its ability to generalize and score completely unseen data in a production environment.

This vignette demonstrates how the `NumericEnsembles` pipeline implements a strict production firewall using the Strength of Concrete data set. We split the dataset externally to simulate an enterprise deployment, pass historical records to the core execution engine, and dispatch predictions onto completely raw holdout data rows using the unified `predict` S3 generic interface.

## 1. Simulating the Production Firewall

We isolate the first 900 rows of the Concrete data to serve as our historical development environment, entirely withholding the remaining 129 rows to simulate unseen production stream inputs.\

```{r, Simulating the Production Firewall}
library(NumericEnsembles)

# Establish the holdout firewall partition
historical_training_data <- Concrete[1:900,]
unseen_production_rows   <- Concrete[901:1030, ]

message(sprintf("Historical training rows: %d", nrow(historical_training_data)))
message(sprintf("Unseen production holdout rows: %d", nrow(unseen_production_rows)))
```

## 2. Model Optimization & Training Domain

We run the core ensemble framework on our historical data to evaluate base architectures, stack meta-learners, and construct blended pairwise options. We will utilize the fast-execution tracking parameters configuration matrix for efficient build cycles.

```{r, Model Optimization and Training Domain}
# Create an explicit configuration matrix block first
custom_config <- NumericEnsemblesConfig(
  cv_folds      = 5,
  train_pct     = 0.70,
  vif_threshold = 5
)

# Pass the config block into the engine safely
Concrete_pipeline <- Numeric(
  dataset       = historical_training_data,
  target_col    = "Strength",
  palette_style = "modern",
  config        = custom_config,
  verbose       = FALSE
)
```

## 3. Dispatched Predictive Inference

We extract the current top-performing leaderboard model variant discovered by the pipeline, and feed the completely raw, un-preprocessed production rows directly into the generic `predict()` method.

```{r, Dispatch Predictive Inference}
# Isolate the champion model name identifier
champion_model <- Concrete_pipeline$performance_report$Model[1]
message(sprintf("Leaderboard Champion discovered: %s", champion_model))

# Generate clean strength projection vectors live
holdout_predictions <- predict(
  object     = Concrete_pipeline,
  newdata    = unseen_production_rows,
  model_name = "best"
)
```

## 4. Evaluation and Performance Audit Summary

Behind the scenes, `NumericEnsembles` automatically tracks performance, identifies structural data patterns, and generates comprehensive audit logs. Here, we compare our live projections side-by-side with ground-truth sales records:

```{r, Evaluation and Performance Audit Summary}
# Synthesize comparison data frame metric results matrix
demo_leaderboard <- data.frame(
  Row_Index       = 901:1030,
  Actual_Strength = unseen_production_rows$Strength,
  Predicted_Strength = round(holdout_predictions, 0),
  Absolute_Error  = round(abs(unseen_production_rows$Strength - holdout_predictions), 0)
)

# Render the first 10 rows of production outputs
knitr::kable(head(demo_leaderboard, 10), row.names = FALSE)

# Compute out-of-sample holdout root mean squared error
final_holdout_rmse <- sqrt(mean((demo_leaderboard$Actual_Strength - demo_leaderboard$Predicted_Strength)^2))
message(sprintf("Final Holdout Production RMSE Score: $%.2f", final_holdout_rmse))
```
